#Savannah Harrison - HW 9
```{r}
library(tidyverse)
library(tidymodels)
bike_data <- read_csv("https://www4.stat.ncsu.edu/~online/datasets/SeoulBikeData.csv",
                      local = locale(encoding = "latin1"))
bike_data
```

```{r}
bike_data |> 
  is.na() |> 
  colSums()
```
```{r}
attributes(bike_data)$spec
```
```{r}
bike_data <- bike_data |>
mutate(date = lubridate::dmy(Date)) |> select(-Date)
```

```{r}
summary(bike_data)
```
```{r}
bike_data$Seasons |> 
  unique()
```
```{r}
bike_data$Holiday |> 
  unique()
```
```{r}
bike_data$`Functioning Day` |> 
  unique()
```
```{r}
bike_data <- bike_data |> 
  mutate(seasons = factor(Seasons),
         holiday = factor(Holiday),
         fn_day = factor(`Functioning Day`)) |> 
  select(-Seasons, -Holiday, -`Functioning Day`)
```

```{r}
bike_data <- bike_data |> rename('bike_count' = `Rented Bike Count`,
         'hour' = "Hour",
         "temp" = `Temperature(°C)`,
         "wind_speed" = `Wind speed (m/s)`,
         "humidity" = `Humidity(%)`,
         "vis" = `Visibility (10m)`,
         "dew_point_temp" = `Dew point temperature(°C)`,
         "solar_radiation" = `Solar Radiation (MJ/m2)`,
         "rainfall" = "Rainfall(mm)",
         "snowfall" = `Snowfall (cm)`)
```

```{r}
bike_data <- bike_data |> 
  filter(fn_day == "Yes") |> 
  select(-fn_day)
```

```{r}
bike_data <- bike_data |>
  group_by(date, seasons, holiday) |> 
  summarize(bike_count = sum(bike_count),
            temp = mean(temp),
            humidity = mean(humidity), 
            wind_speed = mean(wind_speed), 
            vis = mean(vis),
            dew_point_temp = mean(dew_point_temp), 
            solar_radiation = mean(solar_radiation), 
            rainfall = sum(rainfall),
            snowfall = sum(snowfall)) |>
  ungroup()
bike_data
```
```{r}
bike_data |> 
  summarize(across(`bike_count`,
                    .fns = c("mean" = mean, 
                             "median" = median,
                             "sd" = sd,
                             "IQR" = IQR,
                             "min" = min,
                             "max" = max),
                   .names = "{.col}_{.fn}"))
```
```{r}
bike_data |>
  group_by(holiday) |> 
  summarize(across(`bike_count`,
                    .fns = c("mean" = mean, 
                             "median" = median,
                             "sd" = sd,
                             "IQR" = IQR,
                             "min" = min,
                             "max" = max),
                   .names = "{.col}_{.fn}"))
```
```{r}
bike_data |>
  group_by(seasons) |> 
  summarize(across(`bike_count`,
                    .fns = c("mean" = mean, 
                             "median" = median,
                             "sd" = sd,
                             "IQR" = IQR,
                             "min" = min,
                             "max" = max),
                   .names = "{.col}_{.fn}"))
```
```{r}
bike_data |>
  group_by(seasons, holiday) |> 
  summarize(across(`bike_count`,
                    .fns = c("mean" = mean, 
                             "median" = median,
                             "sd" = sd,
                             "IQR" = IQR,
                             "min" = min,
                             "max" = max),
                   .names = "{.col}_{.fn}"))
```
```{r}
bike_data |> 
  select(where(is.numeric)) |> 
  cor() |>
  round(3)
```
```{r}
ggplot(bike_data, aes(x = temp, y = bike_count)) + 
  geom_jitter(aes(color = seasons)) + 
  facet_grid(~holiday)
```
```{r}
ggplot(bike_data, aes(x = solar_radiation, y = bike_count)) + 
  geom_point(aes(color = seasons)) +
  facet_grid(~holiday)
```
```{r}
set.seed(11)
bike_split <- initial_split(bike_data, prop = 0.75, strata = seasons) 
bike_train <- training(bike_split)
bike_test <- testing(bike_split)
bike_10_fold <- vfold_cv(bike_train, 10)
```

```{r}
MLR_rec1 <- recipe(bike_count ~ ., data = bike_train) |>
step_date(date, features = "dow") |>
step_mutate(day_type = factor(if_else(date_dow %in% c("Sat", "Sun"), "Weekend", "Weekday"))) |> step_rm(date, date_dow) |>
step_dummy(seasons, holiday, day_type) |>
step_normalize(all_numeric(), -bike_count)
```

```{r}
MLR_rec2 <- MLR_rec1 |>
step_interact(terms = ~starts_with("seasons")*starts_with("holiday") +
                starts_with("seasons")*temp + 
                temp*rainfall)
```

```{r}
MLR_rec3 <- MLR_rec2 |> 
  step_poly(temp,
            wind_speed,
            vis,
            dew_point_temp,
            solar_radiation,
            rainfall,
            snowfall,
            degree = 2)
```

```{r}
MLR_spec <- linear_reg() |> 
  set_engine("lm")
```

```{r}
MLR_CV_fit1 <- workflow() |> 
  add_recipe(MLR_rec1) |> 
  add_model(MLR_spec) |> 
  fit_resamples(bike_10_fold)
MLR_CV_fit2 <- workflow() |> 
  add_recipe(MLR_rec2) |>
  add_model(MLR_spec) |>
  fit_resamples(bike_10_fold) 
MLR_CV_fit3 <- workflow() |>
add_recipe(MLR_rec3) |> 
  add_model(MLR_spec) |> 
  fit_resamples(bike_10_fold)
```

```{r}
rbind(MLR_CV_fit1 |> collect_metrics(), 
      MLR_CV_fit2 |> collect_metrics(), 
      MLR_CV_fit3 |> collect_metrics())
```
```{r}
final_fit <- workflow() |> 
  add_recipe(MLR_rec1) |> 
  add_model(MLR_spec) |> 
  last_fit(bike_split)
final_fit |> 
  collect_metrics()
```
```{r}
final_fit |> 
  extract_fit_parsnip() |> 
  tidy()
```

#Start of HW9
```{r}
library(tidymodels)
library(caret)
library(glmnet)
library(rpart)
library(randomForest)
library(ipred)
library(ggplot2)
library(recipes)

```

Tuned LASSO model using 10 fold cross-validation.
```{r}
lasso_spec <- linear_reg(penalty = tune()) |>
  set_engine("glmnet") |>
  set_mode("regression")

lasso_grid <- grid_regular(penalty(), levels = 20)

lasso_wf <- workflow() |>
  add_recipe(MLR_rec1) |>
  add_model(lasso_spec)

lasso_tune <- lasso_wf |>
  tune_grid(resamples = bike_10_fold, grid = lasso_grid)
```

Tuned Regression Tree model using 10 fold cross-validation.
```{r}
tree_spec <- decision_tree(cost_complexity = tune(), tree_depth = tune()) |>
  set_engine("rpart") |>
  set_mode("regression")

tree_grid <- grid_regular(cost_complexity(), tree_depth(), levels = 5)

tree_wf <- workflow() |>
  add_recipe(MLR_rec1) |>
  add_model(tree_spec)

tree_tune <- tree_wf |>
  tune_grid(resamples = bike_10_fold, grid = tree_grid)

```

Tuned Bagged Tree model using 10 fold cross-validation.
```{r}
bag_tree_spec <- rand_forest(trees = tune(), min_n = tune()) %>%
  set_engine("randomForest") %>%
  set_mode("regression")

bag_tree_grid <- grid_regular(trees(range = c(50, 500)), min_n(range = c(2, 50)), levels = 5)

bag_tree_workflow <- workflow() %>%
  add_recipe(MLR_rec1) %>%
  add_model(bag_tree_spec)

bag_tree_tune <- bag_tree_workflow %>%
  tune_grid(resamples = bike_10_fold, grid = bag_tree_grid)

```


Tuned Random Forest model using 10 fold cross-validation.
```{r}
rf_spec <- rand_forest(mtry = tune(), trees = tune()) |>
  set_engine("randomForest") |>
  set_mode("regression")

rf_grid <- grid_regular(
  mtry(range =  c(2,10)),
  trees(range = c(100, 200)), 
  levels = 5
)

rf_wf <- workflow() |>
  add_recipe(MLR_rec1) |>
  add_model(rf_spec)

rf_tune <- rf_wf |>
  tune_grid(resamples = bike_10_fold, grid = rf_grid)

```

Finding the best model.
```{r}
lasso_metrics <- collect_metrics(lasso_tune)
tree_metrics <- collect_metrics(tree_tune)
bag_tree_metrics <- collect_metrics(bag_tree_tune)
rf_metrics <- collect_metrics(rf_tune)

lasso_metrics
tree_metrics
bag_tree_metrics
rf_metrics

best_lasso <- lasso_metrics %>%
  filter(.metric == "rmse") %>%
  arrange(mean) %>%
  slice(1)

best_tree <- tree_metrics %>%
  filter(.metric == "rmse") %>%
  arrange(mean) %>%
  slice(1)

best_bag_tree <- bag_tree_metrics %>%
  filter(.metric == "rmse") %>%
  arrange(mean) %>%
  slice(1)

best_rf <- rf_metrics %>%
  filter(.metric == "rmse") %>%
  arrange(mean) %>%
  slice(1)
```

Extracting final model fits for each type.
```{r}
final_lasso <- lasso_wf %>%
  finalize_workflow(best_lasso)

final_tree <- tree_wf %>%
  finalize_workflow(best_tree)

final_bag_tree <- bag_tree_workflow %>%
  finalize_workflow(best_bag_tree)

final_rf <- rf_wf %>%
  finalize_workflow(best_rf)

final_lasso_fit <- final_lasso %>%
  fit(bike_train)

final_tree_fit <- final_tree %>%
  fit(bike_train)

final_bag_tree_fit <- final_bag_tree %>%
  fit(bike_train)

final_rf_fit <- final_rf %>%
  fit(bike_train)
```

Seeing how the models predicts.
```{r}
lasso_pred <- predict(final_lasso_fit, bike_test) %>%
  bind_cols(bike_test) %>%
  metrics(truth = bike_count, estimate = .pred)

tree_pred <- predict(final_tree_fit, bike_test) %>%
  bind_cols(bike_test) %>%
  metrics(truth = bike_count, estimate = .pred)

bag_tree_pred <- predict(final_bag_tree_fit, bike_test) %>%
  bind_cols(bike_test) %>%
  metrics(truth = bike_count, estimate = .pred)

rf_pred <- predict(final_rf_fit, bike_test) %>%
  bind_cols(bike_test) %>%
  metrics(truth = bike_count, estimate = .pred)
```

```{r}
lasso_metrics <- lasso_pred %>%
  filter(.metric %in% c("rmse", "mae"))

tree_metrics <- tree_pred %>%
  filter(.metric %in% c("rmse", "mae"))

bag_tree_metrics <- bag_tree_pred %>%
  filter(.metric %in% c("rmse", "mae"))

rf_metrics <- rf_pred %>%
  filter(.metric %in% c("rmse", "mae"))

list(
  lasso = lasso_metrics,
  tree = tree_metrics,
  bag_tree = bag_tree_metrics,
  rf = rf_metrics
)
```

Reporting final coefficient tables.
```{r}
lasso_coefs <- final_lasso_fit %>%
  extract_fit_parsnip() %>%
  tidy()

lasso_coefs

mlr_coefs <- final_fit %>%
  extract_fit_parsnip() %>%
  tidy()

mlr_coefs
```
Final tree model plot.
```{r}
tree_model <- final_tree_fit %>%
  extract_fit_parsnip()

library(rpart.plot)
rpart.plot(tree_model$fit, type = 4, extra = 101, main = "Final Regression Tree")

```
Creating bagged tree model variable importance plot.
```{r}
bag_tree_model <- final_bag_tree_fit %>%
  extract_fit_parsnip()

library(vip)
vip(bag_tree_model$fit)

```
Creating random forest variable importance plot.
```{r}
rf_model <- final_rf_fit %>%
  extract_fit_parsnip()

vip(rf_model$fit)

```

